---
title: The Ground Game
author: ''
date: ''
slug: []
categories: []
tags: []
---
```{r setup, echo=FALSE, eval=TRUE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
library(tidyverse)
library(dplyr)
library(ggplot2)
library(readr)
library(janitor)
library(stargazer)
require(ggplot2)
require(sf)
library(usmap)
library(rmapshaper)
library(blogdown)
library(Metrics)
```

```{r data, include=FALSE, warning=FALSE}
cvap <- read.csv("cvap_district_2012-2020_clean.csv")
hist <- read.csv("incumb_dist_1948-2020 (3).csv")
expert <- read.csv("expert_rating.csv")

cvap_district <- read_csv("cvap_district_2012-2020_clean.csv")%>%
  rename(st_cd_fips = geoid) 
hist_d<-read_csv("incumb_dist_1948-2020 (3).csv")%>%
  filter(year == 2012 | year == 2014 | year == 2016 | year == 2018 | year == 2020)
expert_d <- read_csv("expert_rating.csv")%>%
  filter(!is.na(district))%>%
  mutate(district_num=as.numeric(district),
         dist_num=case_when(
   district_num<10~paste0('0',district),
    TRUE~district
  ))

dist_cvap <- hist_d %>%
  inner_join(cvap_district, by = c('st_cd_fips', 'year','state'))%>%
  mutate(district_num=as.numeric(district_num))
data <- left_join(dist_cvap, expert_d, by=c('year','state','district_num'))
data <- data %>%
  mutate(totalvotes = RepVotes + DemVotes,
         turnout = totalvotes/cvap,
         DemVotesMajorPct = DemVotesMajorPercent/100,
         RepVotesMajorPct = RepVotesMajorPercent/100)
data_pred <- data %>% filter(!is.na(avg_rating))

data_pred$id <- 1:nrow(data_pred)

train <- data_pred %>% dplyr::sample_frac(0.50)
test  <- dplyr::anti_join(data_pred, train, by = 'id')
```

This week, I examine whether or not there is a relationship between turnout and expert rating. To do so, I run two models: the first seeing if expert rating can predict turnout, and the second doing the same while controlling for incumbency. 

```{r}
lm1<-lm(turnout~avg_rating, data=train)
lm2<-lm(turnout~avg_rating+RepStatus, data=train)

accuracy<-function(model){
  test$pred <- predict(model,test,allow.new.levels=TRUE)
  rmse <- sqrt(mean(test$pred-test$turnout)^2)
  mae <- mae(test$pred,test$turnout)
  return(list('rmse'=rmse,'mae'=mae))
}

stargazer(lm1,lm2,type='text')
```

The first model shows that there is not a strong relationship between the rating and turnout. However, once incumbency is added, there seems to be a stronger relationship, and the data demonstrates that as a a district is predicted to be more Republican, the turnout decreases by about 0.016 points. Though it has more significance, this is still a small value.
